Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2015-12-15 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Revised
Closes #62

\begin{abstract}
Volunteer computing is extensively used in science and industry, but it
still has a long way to go in order to become a mainstream
technology in the same way the cloud technologies in general are. %I have changed "have" to "are"
The lack of general purpose and open source tools with built-in analytics 
could be one of the reasons of why this is so, as well as the lack of
proofs of concept. That is why in this paper we present {\sf NodIO}, a
framework used to create client-server architectures for distributed
evolutionary algorithms, entirely written in JavaScript, which needs
minimal server support to work and that is able to support persistent,
asynchronous, distributed evolutionary algorithms running in the
browser, as well as any other client consuming its REST application
programming interface. An enhanced implementation, {\sf NodIO-W$^2$} is also
presented; this version leverages the HTML5 Web Workers specification
to execute several clients per browser in the background, so that the the number of
generations per user could be improved. The architecture, along with some initial
results on the expected performance and scaling behavior when the
number of users is increased is presented. We also use
the algorithm data to find out what kind of patterns the user behavior
follows by examining the resources they devote to our experiment, and
try to fit them to some statistical distribution to create the bases
of a model of user behavior in volunteer computing experiments of this
kind that can be used by researchers to predict its performance. 
In general, our experiments prove, first, that {\sf NodIO} is a flexible and
low-overhead platform to perform distributed computation experiments
that include volunteers, and then that user behavior follows some patterns,
and, thus, are amenable to a model.
\end{abstract}

>>>>

\begin{abstract}
Volunteer computing is extensively used in science and industry, but
it still has a long way to go in order to become a mainstream
technology in the same way the cloud technologies in general are. %I
have changed "have" to "are" The lack of general purpose and open
source tools with built-in analytics could be one of the reasons of
why this is so, as well as the lack of proofs of concept. That is why
in this paper we present {\sf NodIO}, a framework used to create
client-server architectures for distributed evolutionary algorithms,
entirely written in JavaScript, which needs minimal server support to
work and that is able to support persistent, asynchronous, distributed
evolutionary algorithms running in the browser, as well as any other
client consuming its REST application programming interface. An
enhanced implementation, which we have called {\sf NodIO-W$^2$}, is
also presented; this version leverages the HTML5 Web Workers
specification to execute several clients per browser and do it in the
background, so that the the number of generations per user could be
improved. The architecture, along with some initial results on the
expected performance and scaling behavior when the number of users is
increased is presented. We also use the algorithm data to find out
what kind of patterns the user behavior follows by examining the
resources they devote to our experiment, and try to fit them to some
statistical distribution to create the bases of a model of user
behavior in volunteer computing experiments of this kind that can be
used by researchers to predict its performance.  In general, our
experiments prove, first, that {\sf NodIO} is a flexible and
low-overhead platform to perform distributed computation experiments
that include volunteers, and then that user behavior follows some
patterns, and, thus, are amenable to a model.
\end{abstract}
